### 2020-07-11 Manuscript/sec1-introduction.tex

\section{Introduction}

This is a survey \cite{zhou2018_gnn_review}.