### 2020-07-15 Manuscript/sec5-insights.tex

\section{Insights}

Through our experiments, we prove the existing graph neural network system's performance optimization
technologies. For exampling, \ref{sec:effects_of_sampling_techniques_on_performance} finds that the advantage of 
sampling techniques can greatly reduce the peak memory usage, which clearly explains that when scaling up to large-scale
graph sampling layer is an essential part of AliGraph \cite{zhu2019_aligraph}. \ref{sec:training_time_breakdown} shows that in real graph datasets
(average degress over than 10), collect step of edge calculation is performance bottleneck. DGL \cite{DGL}'s key optimization
methods kernel fusion operation that combines message function with update funsion and requires message function withour parameters.

Compared to \cite{zhang2020_analysis_neugraph}, the autor thinks that computation-intensive GEMM kernel is not the bottleneck of GNN training. This
is not consistent with the hotspot analysis in \ref{sec:training_time_breakdown}. This is because the standard of selecting algorithms of the paper is to cover the varieties for GNN workloads 
while ignoring the complexity of GNN, which is mainly brought by matrix multiplication. In addition, the author's conclusion that kernel fusion can imporove performance
that is similar to our findings.

According to our experiments, we propose the following key findings:

\begin{itemize}
    \item The edge/vertex complexity analysis can be used to theoretically analyze performance bottleneck. And it allows algorithm engineers to use larger hyperparameters to
    increase the complexity of GNN without worrying about the time-consuming training and the explosive growth of memory usage.
    \item Edge computation is performance bottleneck of GNN training for real-world graphs. When the complexity of edge calculation is high, 
    optimization on collect step is most important. When the complexity of edge calculation is low, the collect and aggregate step is the main performance bottleneck.
    \item GPU memory expansion is the most significant fator restricting the scalabiltiy of the GNN algorithm. When dealing with larger-scale graphs, how to effectively save memory is always
    a priority
    \item The advantage of sampling technology is that it greatly reduces the overhead of peak memory usage, making large-scale graph neural network training possible.
    And the time-consuming optimization of sampling technology will also bring great gains to the entire training.
\end{itemize}