### 2020-07-15 Manuscript/sec7-conclusion.tex

\section{Conclusion and Future Work}

We systematically study the performance bottleneck of graph neural networks. We first analyze the GNN training
from the perspective of edge/vertex time complexity and prove the validity of computation analysis on training
time and memory usage. We then decompose the training time from top to bottom and analyze factors that affected GPU memory and
the impact of sampling technology. Our analysis suggests that the edge calculation is the performance bottleneck of GNN model
and optimizing the collection stage and aggregation stage is always worth doing, but when the complexity of computing is high, the message stage is
the most stage that should be optimized. In addition, the huge memory usage is a big problem of GNN training and is linearly related of the number for edges.
Sampling technology is one of ways to greatly reduce the peak memory usage overhead.

In our work, the aggregator required by the selected algorithm is generally min, max, and mean.
When the aggregate becomes complicated, whether the performance brings further changes can be used as future research work.