### 2020-08-21 Manuscript/ms.tex

%! TEX root = ms.tex
%! TEX program = pdflatex
\documentclass{elsarticle}
\usepackage{geometry}
\geometry{left=3cm, right=3cm, top=2.5cm, bottom=2.5cm} 
\usepackage{lineno,hyperref}
\usepackage{amsmath, amsthm}
\usepackage{relsize}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{subfig}
\modulolinenumbers[5]
\bibliographystyle{elsarticle-num}

\begin{document}

\begin{frontmatter}

	\title{Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View}
	\author{Zhaokang Wang, Yunpan Wang}%, Chunfeng Yuan, Yihua Huang\corref{correspondingauthor}}
	%	\cortext[correspondingauthor]{Corresponding author}
	\ead{\{wangzhaokang, wangyp\}@smail.nju.edu.cn}%}, \{cfyuan, yhuang\}@nju.edu.cn}
	\address{State Key Laboratory for Novel Software Technology, \\Department of Computer Science and Technology, Nanjing University, \\Nanjing 210023, China}

	\begin{abstract}
        Graph neural network(GNN) has become a research hotspot in the field of artificial intelligence due to state-of-the-art performance achieved in many graph-related tasks. At the same time, various graph neural network systems have emerged one after another.
        These systems have used their skills in implementation, but there are very few works to study the performance bottleneck on graph neural network training. In contrast to prior works that present characterization of GCNs and select GNNs cover GNN workloads,
		we select typical algorithms by complexities of edge/vertex calculation. We analyze the performance bottleneck by breaking down training time, study the effects of hyper-parameters and sampling technologies, and find factors of memory usage. 
		We finally put forward some key findings for GNN training and hope our work can help GNN system researchers.
	\end{abstract}

	\begin{keyword}
		graph neural network, performance bottleneck, deep learning, characterization
	\end{keyword}

\end{frontmatter}

\linenumbers

\input{sec1-introduction.tex}
\input{sec2-review-of-graph-neural-network.tex}
\input{sec3-experiment-design.tex}
\input{sec4-experiments.tex}
\input{sec5-insights.tex}
\input{sec6-related-work.tex}
\input{sec7-conclusion.tex}

\bibliography{gnnref.bib}

\end{document}