### 2020-12-26 NEUROC/ms.tex
更新abstract

% !TEX program = pdflatex
\documentclass{elsarticle}
\usepackage{geometry}
\geometry{left=3cm, right=3cm, top=2.5cm, bottom=2.5cm} 
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[colorlinks=true, allcolors=red]{hyperref}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{relsize}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{subfig}
\usepackage{float}
\usepackage{times}
\modulolinenumbers[5]
\bibliographystyle{elsarticle-num}

% Vector
\newcommand{\MyVec}[1]{\boldsymbol{#1}}
% Matrix
\newcommand{\MyMat}[1]{\boldsymbol{#1}}
% Vectors with hat
\newcommand{\hvec}[1]{\hat{\boldsymbol{h}}_{#1}}
% Model parameters
\newcommand{\Param}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{frontmatter}

	\title{Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training and Inference with GPUs}
    \author{Zhaokang Wang, Yunpan Wang, Chunfeng Yuan, Rong Gu$^*$ \corref{correspondingauthor1}, Yihua Huang$^*$ \corref{correspondingauthor2}}
    \cortext[correspondingauthor]{Corresponding authors with equal contribution}
    \ead{\{wangzhaokang, wangyp\}@smail.nju.edu.cn, \{cfyuan, gurong, yhuang\}@nju.edu.cn}
	\address{State Key Laboratory for Novel Software Technology, \\Department of Computer Science and Technology, Nanjing University, \\Nanjing 210023, China}

	\begin{abstract}
		The graph neural network (GNN) has become a popular research area for its state-of-the-art performance in many graph analysis tasks.
        %
		Recently, various graph neural network libraries have emerged.
        %
        They make the development of GNNs convenient, but their performance bottlenecks on large datasets is not well studied.
        %
		In this work, we analyze the performance bottlenecks in GNN training and inference with GPUs empirically.
        % 
		A GNN layer can be decomposed into two parts: the vertex and the edge calculation parts.
        %
		According to their computational complexity, we select four representative GNNs (GCN, GGNN, GAT, GaAN) for evaluation.
        % 
		We breakdown their running time and memory usage, evaluate the effects of hyper-parameters, and assess the efficiency of the sampling techniques.
        %
		The experimental evaluation indicates that the edge-related calculation is the performance bottleneck for most GNNs, dominating the training and inference time and memory usage.
        %
		%Future optimizations can focus on it.
        %
        The sampling techniques are essential for GNN training and inference on big graphs with GPUs, but their current implementation still has room for improvement.
	\end{abstract}

	\begin{keyword}
		graph neural network, performance bottleneck analysis, empirical evaluation, machine learning system, GPU
	\end{keyword}

\end{frontmatter}

\linenumbers

\input{sec1-introduction.tex}
\input{sec2-review-of-graph-neural-network.tex}
\input{sec3-experiment-design.tex}
\input{sec4-experiments.tex}
\input{sec5-insights.tex}
\input{sec6-related-work.tex}
\input{sec7-conclusion.tex}

%\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
\bibliography{gnnref.bib}

\end{document}